\texttt{disttree}: Distributional Trees and Forests

### 2017-09-13 pkg/disttree/inst/draft/draft.Rnw
additional arguments 'legend' and 'ylim' in plotting functions 'plot_ll' and 'plot_rmse'
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1576 255f4249-2263-4417-b1ea-8498779e27a7

## function to estimate standard deviation of randomForest for a new observation
# (in randomForest the argument 'keep.inbag' must be set to TRUE)
rf_getsd <- function(object, newobs, rfdata){

  # get predictions for the new observations from all trees
  pred.newobs <- predict(object, predict.all = TRUE, newdata = newobs)

  # vector where the standard deviations from all trees are stored
  sd_trees <- numeric(length = object$ntree)

  # loop over all trees of the forest
  for(i in 1:object$ntree){

    # get data used to build this tree
    obsid <- rep.int(c(1:NROW(rfdata)), as.vector(object$inbag[,i]))
    obs_tree <- rfdata[obsid,]
    rownames(obs_tree) <- c(1:NROW(obs_tree))
    # get predictions for this data from this tree
    pred.obs_tree <- predict(object, newdata = obs_tree, predict.all = TRUE)$individual[,i]

@@

    sd_trees[i] <- sd(obs_node$accel)
  }

  # average of sd over all trees
  sd_newobs <- mean(sd_trees, na.rm = TRUE)

  return(sd_newobs)   
}

@@

## simulation plot functions
# plot RMSE
plot_rmse <- function(simres, type = c("exp", "par")){

@@

    colnames(rmse) <- colnames
    ylim <- c(min(rmse), max(rmse))

@@

    for(i in 1:(length(col)-1)){
      lines(x = simres$x.axis, y = rmse[,(2*i+1)], type = "l", col = col[i+1])
      lines(x = simres$x.axis, y = rmse[,(2*i+2)], type = "l", lty = 2, col = col[i+1])
    }
    legend('topleft', legendnames, 
           col = col, lty = 1, cex = 0.7)
  }

@@

    for(i in 1:(length(col)-1)){
      lines(x = simres$x.axis, y = rmse[,(2*i+1)], type = "l", col = col[i+1])
      lines(x = simres$x.axis, y = rmse[,(2*i+2)], type = "l", lty = 2, col = col[i+1])
    }
    legend('topleft', legendnames, 
           col = col, lty = 1, cex = 0.7)
  }
}

# plot loglikelihood
plot_ll <- function(simres){

@@

    colnames(ll) <- colnames
    ylim <- c(min(ll), max(ll))

@@

    for(i in 2:length(col)){
      lines(x = simres$x.axis, y = ll[,i], type = "l", col = col[i])
    }
    legend('topleft', legendnames, 
           col = col, lty = 1, cex = 0.7)
}
@

@@

  ## get 'fitted.sigma.rf'
  rf_sd <- numeric(length(NROW(mcycle)))

  for(i in 1:NROW(mcycle)){
    rf_sd[i] <- rf_getsd(rf, newobs = mcycle[i,], rfdata = mcycle)
  }

@@

      # get predicted values
      dt.pred.par <- predict(dt, newdata = testdata, type = "parameter")
      df.pred.par <- predict(df, newdata = testdata, type = "parameter")
      rf.pred.resp <- predict(rf, newdata = testdata)
      rf.sd <- numeric(length(NROW(testdata)))
      for(l in 1:NROW(testdata)){
        rf.sd[l] <- rf_getsd(rf, newobs = testdata[l,], rfdata = learndata)
      }
      g.pred.par <- predict(ga, newdata = testdata, type = "response")
      g.pred.mu <- g.pred.par[,1]
      g.pred.sigma <- 1/g.pred.par[,2]

>>>>

  ## function to estimate standard deviation of randomForest for a new observation
  # (in randomForest the argument 'keep.inbag' must be set to TRUE)
  rf_getsd <- function(rf, newdata, rfdata){

    rf_sd <- numeric(length = NROW(newdata))

    for(k in 1:NROW(newdata)){
      newobs <- newdata[k,]
      # get predictions for the new observations from all trees
      pred.newobs <- predict(rf, predict.all = TRUE, newdata = newobs)

      # vector where the standard deviations from all trees are stored
      sd_trees <- numeric(length = rf$ntree)

      # loop over all trees of the forest
      for(i in 1:rf$ntree){

        # get data used to build this tree
        obsid <- rep.int(c(1:NROW(rfdata)), as.vector(rf$inbag[,i]))
        obs_tree <- rfdata[obsid,]
        rownames(obs_tree) <- c(1:NROW(obs_tree))
        # get predictions for this data from this tree
        pred.obs_tree <- predict(rf, newdata = obs_tree, predict.all = TRUE)$individual[,i]

@@

        sd_trees[i] <- sd(obs_node$y)
      }

      # average of sd over all trees
      sd_newobs <- mean(sd_trees, na.rm = TRUE)
      rf_sd[k] <- sd_newobs
    }
    return(rf_sd)   
  }

@@

## simulation plot functions
# plot RMSE
plot_rmse <- function(simres, type = c("exp", "par"), ylim = NULL, legend = TRUE){

@@

    colnames(rmse) <- colnames
    if(is.null(ylim)) ylim <- c(min(na.omit(rmse)), max(na.omit(rmse)))

@@

    for(i in 1:(length(col)-1)){
      lines(x = simres$x.axis, y = rmse[,(2*i+1)], type = "l", col = col[i+1])
      lines(x = simres$x.axis, y = rmse[,(2*i+2)], type = "l", lty = 2, col = col[i+1])
    }
    if(legend) legend('topleft', legendnames, 
           col = col, lty = 1, cex = 0.7)
  }

@@

    for(i in 1:(length(col)-1)){
      lines(x = simres$x.axis, y = rmse[,(2*i+1)], type = "l", col = col[i+1])
      lines(x = simres$x.axis, y = rmse[,(2*i+2)], type = "l", lty = 2, col = col[i+1])
    }
    if(legend) legend('topleft', legendnames, 
           col = col, lty = 1, cex = 0.7)
  }
}

# plot loglikelihood
plot_ll <- function(simres, ylim = NULL, legend = TRUE){

@@

    colnames(ll) <- colnames
    if(is.null(ylim)) ylim <- c(min(na.omit(ll)), max(na.omit(ll)))

@@

    for(i in 2:length(col)){
      lines(x = simres$x.axis, y = ll[,i], type = "l", col = col[i])
    }
    if(legend) legend('topleft', legendnames, 
           col = col, lty = 1, cex = 0.7)
}
@

@@

  ## get 'fitted.sigma.rf'
  rf_sd <- rf_getsd(rf, newdata = mcycle, rfdata = mcycle)

@@

      # get predicted values
      dt.pred.par <- predict(dt, newdata = testdata, type = "parameter")
      df.pred.par <- predict(df, newdata = testdata, type = "parameter")
      rf.pred.resp <- predict(rf, newdata = testdata)

      rf.sd <- rf_getsd(rf, newdata = testdata, rfdata = learndata)

      g.pred.par <- predict(ga, newdata = testdata, type = "response")
      g.pred.mu <- g.pred.par[,1]
      g.pred.sigma <- 1/g.pred.par[,2]
